---
title: "Hands-on_Ex02"
---

# Getting Started

## Install and loading required libraries

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

## Importing data

```{r}
exam_data <- read_csv("data/Exam_data_2.csv")
```

# Working with ggrepel

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## Alternate code to get rid of annotation

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

# Beyond ggplot2 theme

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_minimal() +
  ggtitle("Distribution of Maths scores") 
```

There are 8 built-in themes of ggplot2, which are: theme_gray(), theme_bw(), theme_classic(), theme_dark(), theme_light(), theme_linedraw(), theme_minimal(), and theme_void()

## Working with ggtheme package

ggtheme offers an even wider variety of plotting styles, such as The Economist (theme_economist) or Stata (theme_stata)

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

## Working with hrbthems package

Working with hrbthemes package. Why use this package? It has 2 goals. First, it provides a base theme that focuses on typographic elements. Second, it centers around productivity for a production workflow

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 10,
              base_size = 10,
              grid = "Y")
```

# Beyond single graphs

Graphing of Math scores

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
print(p1)
```

Graphing of English scores

```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
print(p2)
```

Graphing scatterplot for English score versus Maths score

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
print(p3)
```

## Using patchwork extension to create composite figures by combining graphs

install.packages("devtools")

```{r}
library(patchwork)
```

### Combining 2 ggplot2 graphs

```{r}
p1+p2
```

#### Additional attempt 1 : use different code to combine the 2 graphs

```{r}
wrap_plots(p1,p2)
```

### Combining 3 ggplot2 graphs

```{r}
(p1 / p2) | p3
```

#### Additional attempt 2: Trying different style of stacking

```{r}
(p1 | p2) / p3
```

### Creating a composite figure with tag

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

#### Additional attempt 3: Trying to make it more beautiful

```{r}
(p1 | p2) / p3 +
  plot_annotation(tag_levels = 'I') & 
  theme(plot.tag = element_text(size = 10))
```

### Creating figure with insert

```{r}
p3 + inset_element(p2, 
                   left = 0.01, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

#### Additional attempt 4:

```{r}
p2_density <- ggplot(exam_data, aes(x = ENGLISH)) +
  geom_density(fill = "#EEDFCC", alpha = 0.5) +
  ggtitle("English score density") + 
  theme_minimal(base_size = 8) +
  theme(
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    #plot.title = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.3)
  )
print (p2_density)


p3 + inset_element(
  p2_density,
  left = 0.01, bottom = 0.7,
  right = 0.5, top = 1
)

```

### Creating a composite figure by using patchwork and ggtheme combined

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```

#### Additional attempt 5a:

```{r}
patchwork <- (p1 | p2) / p3
patchwork & theme_bw()
```

#### Additional attempt 5b:

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_bw() + theme(plot.title = element_text(size=8))
```

#### Suggested solution:

```{r, fig.width=12, fig.height=7}
patchwork <- (p1 / p2) | p3 
patchwork & theme_bw()
```
